P4IRS: An Intermediate Representation and Compiler for Parallel and Performance-Portable Particle Simulations

### 2024-01-07 manuscript.tex
Update materials and methods
Signed-off-by: Rafael Ravedutti <rafael.r.ravedutti@fau.de>

In this paper we conducted experiments in FAU's clusters Fritz (CPU) and Alex (GPU), these systems have the following configurations:

\begin{itemize}
  \item \textbf{Fritz:} Each compute node has two Intel Xeon Platinum 8360Y “Ice Lake” processors (36 cores per chip) running at a base frequency of 2.4 GHz and 54 MB Shared L3 cache per chip, 256 GB of DDR4 RAM.
  \item \textbf{Alex:} Each compute node has two AMD EPYC 7713 “Milan” processors (64 cores per chip) running at 2.0 GHz with 256 MB Shared L3Cache per chip, 512 GB of DDR4-RAM, eight GPUs are available per node which can be either Nvidia A40 (with 48 GB DDR6 @ 696 GB/s; 37.42 TFlop/s in FP32) or A100 (each 80 GB HBM2 @ 2,039 GB/s; HGX board with NVLink; 9.7 TFlop/s in FP64 or 19.5 TFlop/s in FP32)
\end{itemize}

We perform single-core CPU experiments in both clusters, and refer to both CPUs listed above as Ice Lake and Milan, respectively.
For single GPU experiments, we use both A40 and A100 Nvidia GPUs.
The icc (ICC) 2021.10.0 20230609 compiler was used for generating the CPU code, with the \emph{-Ofast -xCORE-AVX512 -qopt-zmm-usage=high} flags for Ice Lake (AVX512) and \emph{-Ofast -march=core-avx2} flags for Milan (AVX2).
For CUDA, we use the NVCC compiler v12.1.r12.1/compiler.32688072, with same flags as the ones used for the Milan system.
We make use of the Likwid toolkit V5.3 \cite{likwid} to measure HPM counters in the target CPU (Ice Lake) for the single core experiments, where P4IRS is able to properly generate the instrumentation markers and MD-Bench kernels are already instrumented.

>>>>

In this paper we conducted experiments to evaluate the base performance of P4IRS generated codes in the following processors and accelerators:

\begin{itemize}
  \item \textbf{Ice Lake:} Intel Xeon Platinum 8360Y “Ice Lake” processors (36 cores per chip) running at a base frequency of 2.4 GHz and 54 MB Shared L3 cache per chip, 256 GB of DDR4 RAM.
  \item \textbf{Milan:} AMD EPYC 7713 “Milan” processors (64 cores per chip) running at 2.0 GHz with 256 MB Shared L3Cache per chip
  \item \textbf{A40:} Nvidia A40 GPU with 48 GB DDR6 @ 696 GB/s; 37.42 TFlop/s in FP32.
  \item \textbf{A100:} Nvidia A100 GPU with 40 GB HBM2 @ 1,555 GB/s; HGX board with NVLink; 9.7 TFlop/s in FP64 or 19.5 TFlop/s in FP32.
\end{itemize}

And we also perform experiments to evaluate P4IRS weak-scalability in the following clusters:

\begin{itemize}
  \item \textbf{Fritz:} Each compute node has two Ice Lake processors; 256 GB of DDR4 RAM; Blocking HDR100 Infiniband with up to 100 GBit/s bandwidth per link and direction.
  \item \textbf{JUWELS Booster:} Each compute node has two AMD EPYC "Rome" 7402 CPU, with 48 cores (simultaneous multithreading) at 2.8 GHz; 512 GB of DDR4 memory at 3200 MHz; Four Nvidia A100 GPU with 40 GB HBM2e memory; Four InfiniBand HDR (Connect-X6) interconnect channels.
\end{itemize}

The icc (ICC) 2021.10.0 20230609 compiler was used for generating the CPU code, with the \emph{-Ofast -xCORE-AVX512 -qopt-zmm-usage=high} flags for Ice Lake (AVX512) and \emph{-Ofast -march=core-avx2} flags for Milan (AVX2).
For CUDA, we use the NVCC compiler v12.1.r12.1/compiler.32688072, with same flags as the ones used for the Milan system.
We make use of the Likwid toolkit V5.3 \cite{likwid} to measure HPM counters in the target CPU (Ice Lake) for the single core experiments, where P4IRS is able to properly generate the instrumentation markers and MD-Bench kernels are already instrumented.
